---
title: "Areal Spatio-Temporal Kriging"
date: "May 7, 2015"
output: html_document
---

```{r load-data,echo=FALSE, cache=TRUE}
setwd("/Users/xiaomuliu/CrimeProject/SpatioTemporalModeling/ExploratoryAnalysis/CrimeDataSTAnalysis/")
source("importCrimeData.R")
filePath <- "/Users/xiaomuliu/CrimeProject/SpatioTemporalModeling/ExploratoryAnalysis/CPD_DWH/"
fileName <- "X_VIOLENTCRIME_POINTS_01_14.csv"
CrimeData <- importCrimeData(filePath,fileName)
```

In this part we generalize the 2D (space-only) Kriging work to 3D (space-time) Kriging. Note: for this analysis, the mismatch issue between the polygon shape file and the tabular data was resolved by removing all the mismatched records.

```{r conn-mat, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
library(rgdal)
shapefilePath <- "/Users/xiaomuliu/CrimeProject/SpatioTemporalModeling/CPDShapeFiles/"
beat.rg <- readOGR(shapefilePath, "beat_bndy")

# centroids 
Crd.beat <- coordinates(beat.rg)
```

```{r aggregation, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
## **************** Create spatial polygon data frame *****************
# aggregated by "beat"
CrimeData.beat_day <- aggregate(INC_CNT~BEAT+DATEOCC+YEAR+MONTH+DOW,data=CrimeData, FUN=sum, na.rm=TRUE)
CrimeData.beat_day <- CrimeData.beat_day[order(CrimeData.beat_day$DATEOCC),]
CrimeData.beat_day$DOW <- factor(CrimeData.beat_day$DOW, levels=c("Sun","Mon","Tue","Wed","Thu","Fri","Sat"))
CrimeData.beat_month <- aggregate(INC_CNT~BEAT+MONTH,data=CrimeData, FUN=sum, na.rm=TRUE)
CrimeData.beat_year <- aggregate(INC_CNT~BEAT+YEAR, data=CrimeData, FUN=sum, na.rm=TRUE)
CrimeData.beat_all <- aggregate(INC_CNT~BEAT,data=CrimeData, FUN=sum, na.rm=TRUE)

# NOTE: there are a few mismatches between beat number in shape files and those in crime data

# create a spatial polygon data frame template
beat_template.spdf <- beat.rg
beat_template.spdf@data$BEAT_NUMBE<- as.factor(as.integer(as.character(beat_template.spdf@data$BEAT_NUMBE)))
beat_template.spdf@data$INC_CNT <- rep(NA,nrow(beat_template.spdf@data))

## *************** Create Spatio-Temporal irregular layout data ******************
library(spacetime)

# Currently resolve the mismatch issue by removing all the mismatched records 
bNo1 <- as.numeric(sort(as.character(unique(beat.rg@data$BEAT_NUMBE))))
bNo2 <- as.numeric(as.character(unique(CrimeData.beat_all$BEAT)))
miss <- numeric()
for (i in bNo2){
  if (length(which(bNo1==i)) == 0){
    miss <- c(miss,i)
  }
}

rm_idx <- numeric()
for (i in miss){
  k = which(as.numeric(as.character(CrimeData.beat_day$BEAT)) == i)
  rm_idx = c(rm_idx,k)
}
CrimeData.beat_day<-CrimeData.beat_day[-rm_idx,]
CrimeData.beat_day$BEAT <- as.factor(CrimeData.beat_day$BEAT)

CrimeData.beat_day$BEAT <- factor(as.character(CrimeData.beat_day$BEAT),levels=levels(beat_template.spdf@data$BEAT_NUMBE))

beat.loc <- data.frame(BEAT=beat_template.spdf@data$BEAT_NUMBE, X_COORD=Crd.beat[,1], Y_COORD=Crd.beat[,2])
beat.loc <- beat.loc[match(CrimeData.beat_day$BEAT,beat.loc$BEAT),]

beatCentroids <- coordinates(beat.loc[,c("X_COORD","Y_COORD")])
rownames(beatCentroids) <- beat.loc$BEAT
beatCentroids <- SpatialPoints(beatCentroids,CRS(proj4string(beat_template.spdf)))

CrimeData_beat_day.stfdf <- stConstruct(subset(CrimeData.beat_day,select=c("BEAT","DATEOCC","INC_CNT")),space="BEAT",
                                 SpatialObj=beatCentroids, time="DATEOCC")
CrimeData_beat_day.stfdf <- as(CrimeData_beat_day.stfdf,"STFDF") 

```

Before plotting the spatio-temporal variograms, we first look at the autocorrelation function (ACF) of the time series only. Due to the large number of beats, here we only display district-wise ACFs. Note: district 31 does not have adequate data to calculate ACF. The purpose of doing this is to explore if time serie autocorrelations vary according to their locations. Even de-trending process has not been done, we see that the the daily crime counts in a few districts (1, 13, 20, 21) are not very correlated in terms of time. 

```{r acf, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=8, fig.height=6, cache=TRUE}
## ***************** Temporal autocorrelation *************
# plot ACF of each district
CrimeData.district <- aggregate(INC_CNT~DISTRICT+DATEOCC+YEAR+MONTH+DOW,data=CrimeData, FUN=sum, na.rm=TRUE)

# par(mfrow=c(4,3),oma=c(0,0,2,0))
par(mfrow=c(4,3))
for (i in unique(CrimeData.district$DISTRICT)){
  if (i== 31) next #district 31 does not have adequate data
  par(mar=c(4,4,3,1.5))
  acf(na.omit(subset(CrimeData.district,DISTRICT==i,select="INC_CNT")), main = i,lwd=1.5)
}
# title(main=list("Autocorrelation for each district (2008-2014)", cex=1.25),outer=TRUE)

```

For the following analysis, we will be focusing only on beats. Now combining with the spatial covariance information, we plot the city-wide spatio-temporal variogram. For computational reasons, here we only pooled over data of 08/01/2013-07/31/2014 and considered time lag up to 14 days. The spatio-temporal variogram can be viewed in different ways. Two 2D plots and a 3D plot are shown in the following three figures.

```{r stVgm-sample, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=7.5, fig.height=5.5, cache=TRUE,eval=TRUE}
## ********************* spatio-temporal variograms *************************
library(gstat)
timeslot <- "2013-08-01::2014-07-31"
ptm <- proc.time()
stVgm <- variogramST(INC_CNT~1,CrimeData_beat_day.stfdf[,timeslot,"INC_CNT"],tlags=0:14,progress=F)
proc.time() - ptm

# eps = 1e-6
# CrimeData_beat_day.stfdf@data$logINC_CNT <- log(CrimeData_beat_day.stfdf@data$INC_CNT+eps)
# ptm <- proc.time()
# stVgm <- variogramST(logINC_CNT~1,CrimeData_beat_day.stfdf[,"2014-09-01::2014-10-31","logINC_CNT"],tlags=0:14,progress=F)
# proc.time() - ptm

# # reduce spatial separation distance up to which point pairs are included in semivariance estimates
# ptm <- proc.time()
# stVgm <- variogramST(INC_CNT~1,CrimeData_beat_day.stfdf[,"2014-01-01::2014-12-31","INC_CNT"],cutoff=2e4,tlags=0:14,progress=F)
# proc.time() - ptm

f1 <- plot(stVgm,colorkey=list(width=0.75))
f2 <- plot(stVgm,map=FALSE)
# print(f1, position=c(0, 0, 0.5, 1), more=TRUE)
# print(f2, position=c(0.5, 0, 1, 1))
print(f1)
print(f2)

f3 <- plot(stVgm, wireframe=T,zlab=NULL,xlab=list("distance", rot=30), ylab=list("time lag", rot=-35),
     scales=list(arrows=F,z=list(distance=5)), col.regions=bpy.colors(),colorkey=list(width=0.75))
print(f3)
```

Here we are trying out three spatio-temporal variogram models for fitting sample S-T variograms: 

the metric model:
$$
C_{st}(h_s,h_t) = C(a^2|h_s|^2+b^2h_t^2)
$$
the separable (product) model: 
$$
C_{st}(h_s,h_t) = C_s(h_s)C_t(h_t)
$$
the product-sum model:
$$
C_{st}(h_s,h_t) = k_1C_s(h_s)C_t(h_t)+k_2C_s(h_s)+k_3C_t(h_t)
$$
Utilizing the relation $\gamma(h)=C(0)-C(h)$, we can write these covariance expressions in terms of variogram, for example,

the separable model:
$$
\gamma_{st}(h_s,h_t) = C_t(0)\gamma_s(h_s)+C_s(0)\gamma_t(h_t)-\gamma_s(h_s)\gamma_t(h_t)
$$
the product-sum model:
$$
\gamma_{st}(h_s,h_t) = [k_2+k_1C_t(0)]\gamma_s(h_s)+[k_3+k_1C_s(0)]\gamma_t(h_t)-k_1\gamma_s(h_s)\gamma_t(h_t)
$$

For the metric model, we set the spatio-temporal ratio which is defined as the number of space units equivalent to one time unit to be 2500, that is the variogram values of two samples with distance 2500 in space is equivalent to those with time diffence of one day.

```{r stVgm-fit, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=7.5, fig.height=5.5, cache=TRUE, eval=TRUE}
NumBeat <- length(unique(beat_template.spdf@data$BEAT))
stRatio <- 2500
methods = c("metric","separable","productSum")

vgm1 <- vgmST(methods[1],joint=vgm(psill=1,"Sph",range=6000, nugget=0.1),stAni=stRatio)
vgm1.fit <- fit.StVariogram(stVgm,vgm1)
# attr(vgm1.fit,"optim")$value #rmse

vgm2 <- vgmST(methods[2],space=vgm(psill=1,"Sph",range=6000, nugget=0.2),time=vgm(psill=0.3,"Sph",range=14,nugget=0.1), sill=1.2)
vgm2.fit <- fit.StVariogram(stVgm,vgm2,method="L-BFGS-B")
# attr(vgm2.fit,"optim")$value #rmse

vgm3 <- vgmST(methods[3],space=vgm(psill=1,"Sph",range=6000, nugget=0.2),time=vgm(psill=0.3,"Sph",range=14,nugget=0.1), sill=1.2, nugget=0.1)
vgm3.fit <- fit.StVariogram(stVgm,vgm3,method="L-BFGS-B")
# attr(vgm3.fit,"optim")$value #rmse

plot(stVgm, list(vgm1.fit,vgm2.fit,vgm3.fit), colorkey=list(width=0.75))

library(lattice)
# plot(stVgm,vgm1.fit,all=T,wireframe=T,zlab=NULL,xlab=list("distance", rot=30), ylab=list("time lag", rot=-35),
#      scales=list(arrows=F,z=list(distance=5)))
# 
# plot(stVgm,vgm2.fit,all=T,wireframe=T,zlab=NULL,xlab=list("distance", rot=30), ylab=list("time lag", rot=-35),
#      scales=list(arrows=F,z=list(distance=5)))

plot(stVgm,list(vgm1.fit,vgm2.fit,vgm3.fit), all=T, wireframe=T, zlab=NULL, xlab=list("distance", rot=30), 
     ylab=list("time lag", rot=-35),scales=list(arrows=F,z=list(distance=5)), colorkey=list(width=0.75))

```

Now we would like to predict next two day's violent crime distribution. The spatial interpolation was done over a $100 \times 100$ regular grid within the city border. Again, local 200 neigboring observations were used for computational concerns. To be more precisely, the variogram model was estimated globally while solving the linear system equations only take the local observations into account to calucate distances in kernels. We pick the product-sum model which yields the Kriging prediction shown as follows.

```{r STkrige, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=7.5, fig.height=5.5, cache=TRUE, eval=TRUE}
## ************************* Kriging ********************************
## set up prediction grid

# temporal forecast grid
library(xts)
pred.ndays <- 2
# grd.t <- xts(1:pred.ndays,order.by=seq(max(index(CrimeData_beat_day.stfdf))+1,max(index(CrimeData_beat_day.stfdf))+pred.ndays,length=pred.ndays))
grd.t <- xts(1:pred.ndays,order.by=seq(max(index(CrimeData_beat_day.stfdf[,timeslot]))+1,
                                       max(index(CrimeData_beat_day.stfdf[,timeslot]))+pred.ndays,length=pred.ndays))

# spatial interpolation grid
cellsizeX = 377
cellsizeY = 467
interpx = 100
interpy = 100

grd.s <- spsample(beat.rg, n = interpx*interpy, type="regular")
# grd.s <- spsample(beat.rg, cellsize=c(cellsizeX,cellsizeY),type="regular",offset = c(0.5,0.5))
dimnames(grd.s@coords)[[2]]<-c("X_COORD","Y_COORD")
dimnames(grd.s@bbox)[[1]]<-c("X_COORD","Y_COORD")
grd.s <- SpatialPixels(grd.s)

# match grid projection
# grd.s <- spTransform(grd.s, CRS("+proj=utm +zone=16 +datum=WGS84"))
# CrimeData_beat_day.stfdf <- spTransform(CrimeData_beat_day.stfdf, CRS(proj4string(grd.s)))

# spatio-temporal grid
grd.st <- STF(grd.s,grd.t)

# space complexity prohibited (matrix inversion: O(n^2), n ~= 287*t)
# ptm <- proc.time()
# pred <- krigeST(INC_CNT~1, CrimeData_beat_day.stfdf[,"2014-12-01::2014-12-31","INC_CNT"], grd.st, vgm1)
# proc.time() - ptm
# # pred.stfdf <- STFDF(grd, tgrd, data.frame(INC_CNT = pred))
# pred.stfdf <- pred

# local kriging
ptm <- proc.time()
attr(vgm3, "temporal unit") <- "days"
predST <- krigeST(INC_CNT~1, data=CrimeData_beat_day.stfdf[,timeslot,"INC_CNT"], newdata=grd.st, 
                modelList=vgm3.fit, nmax=200, stAni=stRatio, computeVar=FALSE, progress=FALSE)
proc.time() - ptm

# library(RColorBrewer)
# stplot(predST, col.regions=topo.colors(256), mode="xt")

library(lattice)
INC.krigedST <- as.data.frame(predST)
INC.krigedST$endTime <- NULL
INC.krigedST$sp.ID <- NULL
INC.krigedST$V1 <- NULL
names(INC.krigedST)[1:ncol(INC.krigedST)] <- c("X_COORD","Y_COORD","DATEOCC","Pred")

levelplot(Pred~X_COORD+Y_COORD|DATEOCC, data=INC.krigedST,col.regions=topo.colors(256),
          xlab="X Coordinate",ylab="Y Coordinate",main="Kriging Prediction",colorkey=list(width=0.75))

# krigeST has not implemented the code of computing variance for krigeST.local
# levelplot(sqrt(Var)~X_COORD+Y_COORD|DATEOCC, data=INC.krigedST,col.regions=topo.colors(256),
#           xlab="X Coordinate",ylab="Y Coordinate",main="Prediction Standard Error",colorkey=list(width=0.75))
```

The value differences between these two day predictions.
```{r PredDiff, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center',fig.width=5, fig.height=5, cache=TRUE,eval=TRUE}
t1 <- subset(INC.krigedST,DATEOCC==as.Date("2014-08-01"),select=c("X_COORD","Y_COORD","Pred"))
t2 <- subset(INC.krigedST,DATEOCC==as.Date("2014-08-02"),select=c("X_COORD","Y_COORD","Pred"))
dif <- data.frame(X_COORD = t1$X_COORD,Y_COORD = t1$Y_COORD,PredDiff<-(t2$Pred-t1$Pred))
# sum(abs(dif$PredDiff))
levelplot(PredDiff~X_COORD+Y_COORD, data=dif,col.regions=topo.colors(256),
          xlab="X Coordinate",ylab="Y Coordinate",main="Prediction Difference",colorkey=list(width=0.75))
```


